Red Wine Qulaity Exploration by Prateek
========================================================
## Loading essential packages

```{r message=FALSE, warning=FALSE, packages}
library(gridExtra)
library(ggplot2)
library(reshape2)
```

## Loading data set and checking the snapshot of data varibale and values.

```{r Load_the_Data}
winedata = read.csv("C:/Users/prateek/Desktop/wineQualityReds.csv")
head(winedata)
```

# Statiscal Summary
## Structure of data

```{r Univariate_Plots}
str(winedata)
```

## Statistical summary of the variables

```{r}
summary(winedata)
```

* There are **1,599** observation with total **13** variables.
* All varibales are numeric except ```Quality``` and ```X```.
* We can get an idea form ``` quality``` varibale that minimum wine quality is 3
and maximum wine qualtity is 8 with a mean value of 5.6.

# Univariate Plots Section

Looking at the distribution of each varibale in my dataset. Here i have looped 
through each varibale and createda hostograms to look at the distributions of 
the indiviual varibales.

```{r}
winedatap = winedata[-1]
layout(matrix(c(1,2,3,4,5,6,7,8,9,10,11,12),ncol = 4, nrow =3, byrow = TRUE))
for(i in colnames(winedatap)){
  hist(winedatap[[i]], xlab = i, main = i)}
```

# Univariate Analysis

## Creating New Factor Variable

As we can see the entire wine quality are in the range of 3 to 8 with the most 
common values are 5 and 6 and the least common values are 3, 4, 7, and 8. 
So, we create another variable classifing good qualtiy and premium quality wine.

* Good Qulaity Wine = 0
* Premium Quality Wine = 1 

```{r}

winedatap$winequality = ifelse(winedatap$quality > 6, "1", "0")
winedatap$winequality = as.numeric(winedatap$winequality)
hist(winedatap$winequality, xlab = "Quality", main = "Winequalty")

```


* ```Fixed.acidity```, ```volatile.acidity```, ```sulfur.dioxide```, ```sulphates``` 
   and ```alcohol``` are appeared to be long tailed.
* ```Good Quality Wine``` = 1400
* ```Premium Qaultiy Wine``` = 199


## Checking outliers with the help of box plot

```{r}
layout(matrix(c(1,2,3,4,5,6,7,8,9,10,11,12),ncol = 4, nrow =3, byrow = TRUE))
for(i in colnames(winedatap)){
  boxplot(winedatap[[i]], horizontal = T, xlab = i)}

```


* ```Density``` and ```pH``` are normally distributed with few outliers.
* ```Residual.sugar``` and ```chlorides``` have extreme outliers.

## Find some patterns in residual.sugar and chlorides after removing some extreme outliers

```{r, warning=FALSE,message=FALSE}
par(mfrow = c(2,2))
ggplot(data = winedatap, aes(x = residual.sugar)) +
  geom_histogram() +
  scale_x_continuous(lim = c(0.5, quantile(winedatap$residual.sugar, 0.95))) +
  xlab('residual.sugar (g / dm^3)')

```

```{r, warning=FALSE, message=FALSE}
ggplot(data = winedatap, aes(x = chlorides)) +
  geom_histogram() +
  scale_x_continuous(lim = c(0.04, quantile(winedatap$chlorides, 0.95))) +
  xlab('chlorides (g / dm^3)')
```


* After removing some extreme outliers, we see chlorides are now normally distributed



## Check diffirence in mean value for the overall distribution and distribution for premium qulaity wine.

If the diffirence in too high we can get a small idea if that particular varibale helps in creating a premium wine.

```{r}
goodqualitydata = winedatap[winedatap$quality>6,]
for(i in colnames(winedatap)){
  for(j in colnames(goodqualitydata)){
    a = mean(goodqualitydata[[j]])}
  b = mean(winedatap[[i]])
  print(i)
  print(b-a)
}

```


* ``` total.sulfur.dioxide``` and ``` free.sulfur.dioxide``` has a very high 
mean difference.
* ``` volatile.acidity```, ```citiric acid```, ``` chlorides```, ``` density``` 
and ```sulpahtes``` has a considerable mean difference.



# Questions

### What is the structure of your dataset?
There are 1599 items and 14 variables/features in the dataset. There is no 
factor varibale.


### What is/are the main feature(s) of interest in your dataset?

The main feature of interest is quality. I have created one factor variable 
in which i have classified wine qulaity of 7 and 8 as premium quality wine and 
other as good Quality wine.


### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

As of now we by looking at the mean difference, we can 
say ``` total.sulfur.dioxide```,``` free.sulfur.dioxide```,``` volatile.acidity```
, ```citiric acid```, ``` chlorides```, ``` density``` and ```sulpahtes``` can 
help in supporting the investigation.



### Did you create any new variables from existing variables in the dataset?
Yes i have created one varible wineq from qualtiy.There are two levels 
to the varibles , Normal and Premium Wine.


### Of the features you investigated, were there any unusual distributions? \
 
* Residual.sugar and chlorides contains many outliers but after doing some 
operations, chlorides get into normal distribution





# Bivariate Plots Section

## Checking corelation between varibales.

```{r}
cor(winedatap)
```



```{r echo=FALSE, Bivariate_Plots}
winedatap$quality = as.factor(winedatap$quality)
plot(winedatap)
```

* There are very little positive correlation between ```fixed.acidity```, ```citirc.acid```
  and negetive correlation ```fixed.acidity``` and ```Ph```.
  
  There are no two varibales with a very string correlation. But we will 
  check the correlation effects, while creating the logistic regression model.
  
  
# Binning each variable and plotting it with the percentage of premium wine for that bucket.  
  

```{r}
winedatap$quality = as.numeric(winedatap$quality)
layout(matrix(c(1,2,3,4,5,6,7,8,9,10,11,12), ncol = 4, nrow = 3, byrow = TRUE))
for(i in colnames(winedatap)){
  plot.default(aggregate(winedatap$winequality, by = list(cut(winedatap[[i]], 10)), FUN = mean), xlab = i, ylab= "frequency")
  }


```

* We can see here that as th values in the bins increases for ```fixed.acidity```,
and ```alcohol``` the percentage of premium wine increases.
* When the value in ```volatile.acidity``` and ```total.sulphur.dioxide`` 
increses the pecenatge of premium wine decreases.

## Checking the relation between dependent varibale with independent varibales with the help of box plot
### Box Plot of Quality

```{r}

grid.arrange(ggplot(data = winedatap, aes(x = as.factor(quality) , y = fixed.acidity))+geom_boxplot(fill = "green") + xlab("quality") + ylab("fixed acidity (tartaric acid - g / dm^3)"),
             ggplot(data = winedatap, aes(x = as.factor(quality), y= volatile.acidity))+geom_boxplot(fill = "green") + xlab("quality")+ ylab("volatile acidity (acetic acid - g / dm^3)"),
             ggplot(data = winedatap, aes(x = as.factor(quality), y= citric.acid))+geom_boxplot(fill = "green") + xlab("quality")+ ylab(" citric acid (g / dm^3)"),
             ggplot(data = winedatap, aes(x = as.factor(quality), y= residual.sugar))+geom_boxplot(fill = "green") + xlab("quality")+ ylab("residual sugar (g / dm^3)"),
             ggplot(data = winedatap, aes(x = as.factor(quality), y= chlorides))+geom_boxplot(fill = "green")+ xlab("quality")+ ylab("chlorides (sodium chloride - g / dm^3"),
             ggplot(data = winedatap, aes(x = as.factor(quality), y= free.sulfur.dioxide))+geom_boxplot(fill = "green")+ xlab("quality")+ ylab(" free sulfur dioxide (mg / dm^3)"),
             ggplot(data = winedatap, aes(x = as.factor(quality), y= total.sulfur.dioxide))+geom_boxplot(fill = "green") + xlab("quality")+ ylab("total sulfur dioxide (mg / dm^3)"),
             ggplot(data = winedatap, aes(x = as.factor(quality), y= density))+geom_boxplot(fill = "green")+ xlab("quality")+ ylab("density (g / cm^3)"),
             ggplot(data = winedatap, aes(x = as.factor(quality), y= pH))+geom_boxplot(fill = "green")+ xlab("quality")+ ylab("pH"),
             ggplot(data = winedatap, aes(x = as.factor(quality), y= sulphates))+geom_boxplot(fill = "green")+ xlab("quality")+ ylab(" sulphates (potassium sulphate - g / dm3)"),
             ggplot(data = winedatap, aes(x = as.factor(quality), y= alcohol))+geom_boxplot(fill = "green")+ xlab("quality")+ ylab("alcohol (% by volume)"), ncol = 4)

```


### Boxplot of WineQuality

```{r}
grid.arrange(ggplot(data = winedatap, aes(x = as.factor(winequality) , y = fixed.acidity))+geom_boxplot(fill = "green") + xlab("winequality") + ylab("fixed acidity (tartaric acid - g / dm^3)"),
             ggplot(data = winedatap, aes(x = as.factor(winequality), y= volatile.acidity))+geom_boxplot(fill = "green") + xlab("winequality") + ylab("volatile acidity (acetic acid - g / dm^3)"),
             ggplot(data = winedatap, aes(x = as.factor(winequality), y= citric.acid))+geom_boxplot(fill = "green")+ xlab("winequality")+ ylab(" citric acid (g / dm^3)"),
             ggplot(data = winedatap, aes(x = as.factor(winequality), y= residual.sugar))+geom_boxplot(fill = "green")+ xlab("winequality")+ ylab("residual sugar (g / dm^3)"),
             ggplot(data = winedatap, aes(x = as.factor(winequality), y= chlorides))+geom_boxplot(fill = "green")+ xlab("winequality")+ ylab("chlorides (sodium chloride - g / dm^3"),
             ggplot(data = winedatap, aes(x = as.factor(winequality), y= free.sulfur.dioxide))+geom_boxplot(fill = "green")+ xlab("winequality") + ylab(" free sulfur dioxide (mg / dm^3)"),
             ggplot(data = winedatap, aes(x = as.factor(winequality), y= total.sulfur.dioxide))+geom_boxplot(fill = "green")+ xlab("winequality") + ylab("total sulfur dioxide (mg / dm^3)") ,
             ggplot(data = winedatap, aes(x = as.factor(winequality), y= density))+geom_boxplot(fill = "green")+ xlab("winequality") + ylab("density (g / cm^3)"),
             ggplot(data = winedatap, aes(x = as.factor(winequality), y= pH))+geom_boxplot(fill = "green")+ xlab("winequality") + ylab("Ph"),
             ggplot(data = winedatap, aes(x = as.factor(winequality), y= sulphates))+geom_boxplot(fill = "green")+ xlab("winequality") + ylab("sulphates (potassium sulphate - g / dm3)"),
             ggplot(data = winedatap, aes(x = as.factor(winequality), y= alcohol))+geom_boxplot(fill = "green")+ xlab("winequality") + ylab("alcohol (% by volume)"), ncol = 4)

```

* With the help of above plots we can see that ```volatile.acidity```,
```citric.acid```, ```Sulphates``` and ```alcohol``


# Bivariate Analysis

### As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion ? 

I have seen that between independent variables there are varibales with very 
less correlation. Also few variables are highly correalted with decision variable 
that is quality of wine. Out of all variables the varibales with the  highest 
correlation  are ```Volatile.acidity```, ```citric.acid```, ```sulphates``` and
```alcohol```.



### Talk about some of the relationships you observed in this part of the How did the feature(s) of interest vary with other features in the dataset?


Sulphtes and alcohol has a positive correlation whereas volatile.acidity
has negetive correlation with the decsion variable.


### Did you observe any interesting relationships between the other features ?

There is no as such relation between other features. 
Since no features are strongly correlated with each other. 

### What was the strongest relationship you found?

The strongest reltionship is between ```volatile.acidity``` and ```quality of wine```
and ```citric.acid``` and ```quality of wine```.






# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}
winedatap$winequality = as.factor(winedatap$winequality)
ggplot(winedatap, aes(x = volatile.acidity, y = citric.acid,
                      color = winequality))+geom_point() + xlab("Volatile.Acidity((tartaric acid - g / dm^3)") + ylab("citric acid (g / dm^3)")
```

This above graph shows the relation between volatile acidity and citric acid. To attain 
the premium quality wine we can the value of volatilea acidity should have lower value, wehreas 
values of citic acid should be on the higher side.

```{r}
ggplot(winedatap, aes(x = volatile.acidity, y = sulphates, 
                      color = winequality))+geom_point() +  xlab("sulphates (potassium sulphate - g / dm3)") + ylab("volatile acidity (acetic acid - g / dm^3)")
```

Above is the plot between volatile acidity and sulphates where we can see the 
data for premium wine is concentrated on the fourth quandrant. where it is showing
that again for getting premium wine, one should use low value of volatile acidity and 
hig value of sulphates.




```{r}
ggplot(winedatap, aes(x = volatile.acidity, y = alcohol, 
                      color = winequality))+geom_point() + xlab("alcohol (% by volume)") + ylab("volatile acidity (acetic acid - g / dm^3)")
```

This above graphs shows relationship between volatile acidity and alocohol, whcih depicts that 
there should be high alcohol content for remium wines and as usual volatile acidity values should be lower.




```{r}
ggplot(winedatap, aes(x = alcohol, y = sulphates, 
                      color = winequality))+geom_point() + xlab("alcohol (% by volume)") + ylab(" sulphates (potassium sulphate - g / dm3)")

```

This is graph is between aclcohol and sulphates. With the graph we can get an idea
that for premium wines the alcohol content should te on the highers with compartively low values for sulphates.


```{r}
ggplot(winedatap, aes(x = alcohol, y = citric.acid, 
                      color = winequality))+geom_point() + xlab("alcohol (% by volume)") + ylab("citric acid (g / dm^3)")

```

The above graph shown is between alcohol and citric acid, which tells us that
both alcohol and citric acid both should have higher value for creating the 
premium wine.


```{r}
ggplot(winedatap, aes(x = sulphates, y = citric.acid, 
                      color = winequality))+geom_point() + xlab(" sulphates (potassium sulphate - g / dm3)") + ylab("citric acid (g / dm^3)")

```

The last graph is between sulphate and citric acid where it is clearly deicting 
that most of the values of premium wine have higher values of citric but lower 
concetration of sulphate.


# Multivariate Summary

*I have create six multivariate plots, between ```Volatile.acidity```, 
```citric.acid```, ```sulphates``` and```alcohol```. 
* From these six graphs it is clear that to create premium wine we need lower 
concentration of ```volatile acidity``` but higher concentration of ```alcohol```,
```citric acid```and ```sulphates```.



# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the  investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

Yes there are few relations which makes it clear to make premium wines.For Example
if we talk about voaltile acidity and citric acid , we found that if higher 
citric acid and lower volatile acidity creates premium wines. 


### Were there any interesting or surprising interactions between features?

No there is no as such surpising interactions between variables.

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model?

No i have not created any models fo this.



# Final Plots and Summary


### Plot One : Effect of alcohol on wine quality

```{r Plot_One}
ggplot(data = winedatap, aes(as.factor(quality), alcohol, fill = winequality)) +
  geom_boxplot() +
  ggtitle('Alcohol Effect on Wine Quality') +
  xlab('Quality') +
  ylab('Alcohol (% volume)') 

```

## Description One

As alcohol is highly correlated with the quality, it is better to see its pattern with varying rating. From the above plot, it clearly shows higher % of alcohol yields better wine.


### Plot Two : Effect of acids on wine quality
```{r echo=FALSE, Plot_Two}
grid.arrange(ggplot(data = winedatap, aes(as.factor(quality), fixed.acidity, fill = winequality)) +
               geom_boxplot() +
               xlab("Quality") +
               ylab("Fixed Acidity (g / dm^3)") +
              scale_fill_brewer(type = 'seq', palette = 1),
             ggplot(data = winedatap, aes(as.factor(quality), volatile.acidity, fill = winequality)) +
               geom_boxplot() +
               xlab("Quality") +
               ylab("Volatile Acidity (g / dm^3)") +
              scale_fill_brewer(type = 'seq', palette = 1),
             ggplot(data = winedatap, aes(as.factor(quality), citric.acid, fill = winequality)) +
               geom_boxplot() +
               xlab("Quality") +
               ylab("Citric Acid (g / dm^3)") +
              scale_fill_brewer(type = 'seq', palette = 1),
             ncol = 1)
```

## Description Two

As more the acidic better is the wine. It would be better to see which acids 
have more impact on wine quality. Above plot shows, ```fixed.acidity``` and 
```citric.acid``` have highly correlated with quality but ```volatile.acidity```
has negative impact on quality.



### Plot Three : Effect of citric acid and volatile acid on quality of wine.
```{r echo=FALSE, Plot_Three}
winedatap$winequality = as.factor(winedatap$winequality)
ggplot(winedatap, aes(x = volatile.acidity, y = citric.acid,
                      color = winequality))+geom_point() + xlab("Volatile.Acidity((tartaric acid - g / dm^3)") + ylab("citric acid (g / dm^3)")

```

## Description Three

With voaltile acidity and citric acid , we found that if higher 
citric acid and lower volatile acidity creates premium wines.


------

# Reflection

Red Wine data contains 1599 data points for 13 variables. Basically it 
contains the quality of wine for diffirent values of these 13 variables.
These varibales include, Fixed.acidity,Volatile.acidity,citric.acid,Residual.sugar
,Chlorides,Free.sulfur.oxide, Total.sulfur.dioxide,Density,Ph,Sulpahates,alcohol. 


From this data my main aim is to predict the quality of wine on the 
basis of diffirent variable combination. FOr this the first thing what 
i have done is i have analysed distribution and summary of different variables 
individually and checked the NA values and Outliers values.After that i again 
analysed the distribution of each variables for premium wine. 
For ease of analysis i have categorized quality of wine in two parts good and premium. 


Then after that we checked the if there is any variables (leaving decision variable aside), 
are higly correlated to each other or not, and then checked the 
correlation of each variable with varibale if interest. From these action we 
found that there are four varibale which can actually predict the 
quality of wine.They are ```Volatile.acidity```, ```citric.acid```, 
```sulphates``` and ```alcohol```



Some limitation of the model is that it is little bit biased in a way 
that we donot have enough data points for winequality of 5 and 6. I 
would like apply classifier model in the future to predic the qulaity 
of wine on the basis of values of diffirent varibales.










